---
title: "Descriptive Epidemiology Assignment"
author: "PHW253 Teaching Team"
output:
  pdf_document: default
  pdf: default
---

```{r setup, include=FALSE}
# Show R code in output for all R chunks
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

In this R Markdown (Rmd) file we provide a template to help you complete the R assignments for the first three weeks of the course. 

The sections (Week 1, Week 2, and Week 3) build on each other sequentially. So you will need your R code from Week 1 to complete Week 2, and your code from Weeks 1 and 2 to create your final submission in Week 3. We suggest keeping all you code for this 3-part assignment in this one Rmd.

The dataset for this assignment is Descriptive_Epi_Assignment_line_list_data.xlsx.

--------------------------------------------------------------------------------

# Week 1 - Data Check for the Descriptive Epi Assignment

## Step 1: You may have to install new packages for this assignment. 

```{r}
# (just run once and then comment out each install statement by adding # in front of each line so you are not reinstalling the packages every time you run this file.)
install.packages("rio")
install.packages("dplyr")
install.packages("janitor")
install.packages("flextable")
install.packages("lubridate")
install.packages("incidence2")
install.packages("ggplot2")
```

```{r}
# load packages
# run every time

# rio needed for import code
library("rio")

# dplyr needed for mutate or recode
library(dplyr)

# janitor needed for crosstable
library(janitor)

# flextable needed for export formatted table to Word file
library(flextable)

# for dates in R
library(lubridate)

# ggplot2 for epicurves & data visualization
library (ggplot2)

# incidence2 for epicurves of linelist data
library (incidence2)
```

## Step 2: Import the data into the R environment and look at the structure of the dataset (that is, the rows and columns).

Refer to the R Assignments page on bCourses for detailed instructions and suggested R videos to review.

```{r}
# ADD CODE HERE
```

## Step 3: create new dataset and keep only rows where MLVA is an A, A2, A3 pattern. MLVA is an acronym that stands for Multilocus
 Variable-number tandem-repeat Analysis, a molecular typing technique to help distinguish between different strains of E. coli.

* There are 92 observations in the data set, with 75 cases with MLVA pattern A, 1 case with pattern A2, and 1 case with pattern A3.
* In the data set, you will notice there are many MLVA patterns (that is, there are patterns A, A2, and A3).
* You will have to select out only the rows where MLVA is A, A2, or A3.
* The other E. coli cases that did not have MLVA pattern A, A2, or A3 are not part of this outbreak and can be discarded from the analysis.

```{r}
#ADD CODE HERE
```


## Step 4: Focus on the hospitalization variable and the tabulate number of hospitalizations "Yes" "No" and "Unknown"

```{r}
#ADD CODE HERE
```

--------------------------------------------------------------------------------

# Week 2 - Data Check for the Descriptive Epi Assignment

Refer to the R Assignments page on bCourses for detailed instructions and suggested R videos to review.

## Step 1: Focus on the State variable and tabulate the number of observations for each state.
```{r}
#ADD CODE HERE
```

## Step 2: Define region using census definition  https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf
```{r}
#ADD CODE HERE
```

Step 3: Look at region variable and tabulate the number of cases by region
```{r}
#ADD CODE HERE
```

Step 4: Dates in R: What format is OnsetDate in?
```{r}
#ADD CODE HERE
```

Step 5: Create a new variable with onset date in ymd date format and confirm new variable format
```{r}
#ADD CODE HERE
```

Step 7: Tabulate the new variable to obtain date ranges and determine if there are missing values
```{r}
#ADD CODE HERE
```


--------------------------------------------------------------------------------

# Week 3 – Descriptive Epidemiology Assignment Final Submission


For your final submission, we suggest you make a copy of this Rmd and then reformat it to meet the requirements of the final submission, as detailed on the R Assignments page on bCourses.

