---
title: "Hypothesis Testing Assignment"
author: "PHW253 Teaching Team"
output:
  pdf_document: default
  pdf: default
---

```{r setup, include=FALSE}
# Show R code in output for all R chunks
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

In this R Markdown (Rmd) file we provide a template to help you complete the R assignments for the last three weeks of the course. 

The sections (Week 4, Week 5, and Week 6) build on each other sequentially. So you will need your R code from Week 4 to complete Week 5, and your code from Weeks 4 and 5 to create your final submission in Week 6. We suggest keeping all you code for this 3-part assignment in this one Rmd.

The dataset for this assignment is Hypothesis_Testing_Assignment_case_control_data.xlsx

--------------------------------------------------------------------------------

# Week 4 - Data Check for the Hypothesis Testing Assignment

Call in the necessary R packages for this assignment.

```{r}
# install packages
# (just run once and then comment out by adding # in front of line)
install.packages("rio")
install.packages("skimr")
install.packages("dplyr")
install.packages("janitor")
install.packages("survival")
```

## Step 1: Load packages
 
```{r}
# run every time

# rio needed for import code
library(rio)
# dplyr needed for mutate or recode
library(dplyr)
# janitor needed for crosstable
library(janitor)
# survival needed for clogit
library(survival)
```


## Step 2: Import the data. Examine the variable names and use the janitor package to clean variable names.

Refer to the R Assignments page on bCourses for detailed instructions and suggested R videos to review.

In Weeks 1 to 3, we used line listing (a dataset with only the cases). In Weeks 4 to 6, we will use the outbreak investigation dataset (a matched case-control study).

```{r}
# import
# ADD CODE HERE


# look at the dataset
# ADD CODE HERE. example: names(dataset)


# janitor package clean names changes the column names to lower case & replaces
# the spaces in column names with an underscore
# ADD CODE HERE. example: newnamedataset<- dataset  %>% janitor::clean_names()

```

## Step 3: Subset the data (all code is provided).

The data set contains more records than needed for the matched case-control study. Thus, you will have to subset the data in order to get the study data set for the case-control analysis.

```{r}
## Select only cases that have matched controls
ecoli2 <- ecoli1 %>% 
  filter(cdcid !="CDC001" & cdcid !="CDC006" & cdcid !="CDC009" & cdcid !="CDC012" & 
           cdcid !="CDC015" & cdcid !="CDC022" & cdcid !="CDC023" & cdcid !="CDC024" & 
           cdcid !="CDC037" & cdcid !="CDC043" & cdcid !="CDC045" & cdcid !="CDC046" & 
           cdcid !="CDC048" & cdcid !="CDC049" & cdcid !="CDC067" & cdcid !="CDC068")   

## Some cases have multiple controls
ecoli2 %>%  filter(cdcid == "CDC047")  %>% tabyl(case)

## Only keep controlletter=A for cdcid=="CDC047"
x<-which(ecoli2$cdcid =="CDC047" & ecoli2$controlletter=="B")   
ecoli3 <-ecoli2[-x, ]    

dim(ecoli3)
```

## Step 4: Tabulate number of total cases and total controls in the cleaned dataset.

In Week 6, these are the numbers you'll add to the column headings in the provided table template. 

```{r}
# ADD CODE HERE
```

## Step 5: Look at the strawberry variable. For conditional logistic regression, all values need to be 0 or 1. Recode other values to missing. 

```{r}
# ADD CODE HERE
```

## Step 6: Tabulate the count of cases who ate strawberries. Tabulate the count of controls who ate strawberries. 

```{r}
# ADD CODE HERE
```

## Step 7: Calculate the percent exposed among all cases. Calculate the percent exposed among all controls.

```{r}
# ADD CODE HERE
```



--------------------------------------------------------------------------------

# Week 5 - Data Check for the Hypothesis Testing Assignment

Refer to the R Assignments page on bCourses for detailed instructions and suggested R videos to review.

## Step 1: Create a numeric variable that identifies each pair (all code is provided). This is called strata in the regression code. First, select last two characters of the CDCID string/character variable (new variable stratachar). Then, convert to a numeric format (new variable stratanum).

```{r}
ecoli3$stratachar<-substr(ecoli3$cdcid,start=5,stop=6)

ecoli3 %>%  tabyl(stratachar)

ecoli3$stratanum<-as.numeric(ecoli3$stratachar)

ecoli3 %>%  tabyl(stratachar,stratanum)
```

## Step 2: Conditional logistic regression.
 
```{r}
# ADD CODE HERE
```



--------------------------------------------------------------------------------

# Week 6 – Hypothesis Testing Assignment Final Submission


For your final submission, we suggest you make a copy of this Rmd and then reformat it to meet the requirements of the final submission, as detailed on the R Assignments page on bCourses.

